Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf<-read.csv('pseudo_facebook.tsv', sep='\t')
qplot(x=age, y=friend_count, data=pf)
qplot(age, friend_count, data=pf)



```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha=1/20) 
  xlim(13,90)
```

***

### Overplotting
Notes:

```{r Overplotting}
#or (geom_jiiter adds noise)
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(alpha=1/20) +
  xlim(13,90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha=1/20) +
  xlim(13,90) +
  coord_trans(y='sqrt')
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_jitter(alpha=1/10, position=position_jitter(h=0)) +
  coord_trans(y='sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
install.packages('dplyr')
library(dplyr)

#filter()
#group_by()
#mutate()
#arrange()

age_groups <- group_by(pf,age)
pf.fc_by_age <-summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n=n())

pf.fc_by_age <-arrange(pf.fc_by_age, age)
head(pf.fc_by_age)

#or
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
          friend_count_median = median(as.numeric(friend_count)),
          n=n()) %>%
  arrange(age)
 head(pf.fc_by_age) 

```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + 
  geom_line() +
  coord_cartesian(xlim = c(13, 90)) 
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  coord_cartesian(xlim = c(13,70), ylim = c(0,1000)) +
  geom_jitter(alpha=1/10, position=position_jitter(h=0), color='orange') +
  geom_line(stat='summary', fun.y=mean) +
  geom_line(stat='summary', fun.y=quantile,  fun.args = list(prob = 0.1), linetype =2, color = 'blue') +
  geom_line(stat='summary', fun.y=quantile,  fun.args = list(prob = 0.9), linetype =2, color = 'blue') +
  geom_line(stat='summary', fun.y=quantile,  fun.args = list(prob = 0.5), color = 'blue')
  
 
```

#### What are some of your observations of the plot?
Response:
Havinm mpre than 1000 friends is rare even for youn users, and 90th percentile is below 1000

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count, meathd ='pearson')
#or
with(pf, cor.test(pf$age, pf$friend_count, meathd ='pearson'))

# no meaningful correlation
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age<=70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:
```{r}
#Monotonic relationship
with(subset(pf, age<=70), cor.test(age, friend_count, method = 'spearman'))
```

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x= likes_received, y  = www_likes_received), data = pf) +
  geom_jitter(position=position_jitter(h=0)) 

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}


ggplot(aes(x= www_likes_received, y  = likes_received), data = pf) +
  geom_jitter(position=position_jitter(h=0)) +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method ='lm', color='red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf, cor.test(www_likes_received, likes_received))
```

Response:

***

### Moira on Correlation
Notes:
When doing modeling/regression one of the assumptions of the regression is that variables thrown into regression are independent of each other and if variables are highly correlated then it would be diffucult to tel which ones are actually driving the phenomena. It is important to measure correlation between variables first because it will help pyuo determine which ones you do not want to throw in together and which ones to keep.

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x=Month, y=Temp), data=Mitchell) + 
  geom_point()

#or

qplot(data=Mitchell, Month, Temp)
  
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
#weak correlation
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x=Month, y=Temp), data=Mitchell) + 
  scale_x_continuous( breaks = seq(0, 203, 12)) +
  geom_point()
```

***

### A New Perspective
```{r}
ggplot(aes(x=Month, y=Temp), data=Mitchell) + 
  scale_x_continuous( breaks = seq(0, 203, 12)) +
  geom_point()
```



What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + 
  geom_line()
head(pf.fc_by_age,10)
pf.fc_by_age[17:19,]

pf$age_with_months <- pf$age + (12-pf$dob_month)/12
```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n=n()) %>%
  arrange(age_with_months)
 
head(pf.fc_by_age_months) 


#or
age_groups <- group_by(pf,age_with_months)
pf.fc_by_age_months <-summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n=n())

pf.fc_by_age_months <-arrange(pf.fc_by_age_months, age_with_months)
head(pf.fc_by_age_months)



```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months<71)) + 
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <-ggplot(aes(x = age, y = friend_count_mean), data = subset(pf.fc_by_age, age<71)) + 
  geom_line() +
  geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months<71)) + 
  geom_line() +
  geom_smooth()

p3 <- ggplot(aes(x=round(age/5)*5, y=friend_count), data =subset(pf, age<71) ) +
  geom_line(stat='summary', fun.y=mean)

library(gridExtra)
grid.arrange(p2, p1, p3, ncol=1)

#Notes: By decreasing size of bins and increasing number of bins we have less data to estimate each conditional mean, noise is much worse on age with months since we have finer bin chocies

#Notes om geom_smooth(): while th esmoother captures some of the fetures of this relationship it does not draw attention to non motonic relationship in low anges and misses discontinuity at age 69, this highligts that it can be useful but  like any model it can be subject to systematic errors when the true process of generating the data is not so consistent with the model itself. Above models are based on idea that the true function is smooth but we know that there are some discontinuity in relationship


```

***

### Which Plot to Choose?
Notes:  choose the best 1,2 that best communicate findings of your work

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

