Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom='point', shape=4)

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color=gender), stat='summary', fun.y=median)
#median count is larger for women than men for mostly all ages


pf.fc_by_age_gender <- subset(pf, !is.na(gender)) %>%
  #or instead of subset filter(!is,na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n=n()) %>%
  ungroup() %>%
  arrange(age)
  
head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x=age, y=median_friend_count), data=pf.fc_by_age_gender) +
  geom_line(aes(color = gender))

```

***

### Thinking in Ratios
Notes: Gender difference is greater for young ages.

***

### Wide and Long Format
Notes:Long format (many rows, variables we grouped by (male female) are repeated for each year)

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
#dcast for tesult to be a dataframe, acast for array/matrix
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var='median_friend_count')
#or
pf.fc_by_age_gender.wide <-     subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],                   !is.na(gender)) %>% 
    spread(gender, median_friend_count) %>% 
    mutate(ratio = male / female) 

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x=age,y=female/male), data=pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(yintercept=1,linetype=2, alpha=0.3)
#For very young users the median female user has over 2.5 times as many friends as median male user 
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)

table(pf$year_joined)
pf$year_joined.bucket <-cut(pf$year_joined, c(2004,2009, 2011, 2012, 2014))
table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x=age, y=friend_count), data=subset(pf, !is.na(year_joined))) +
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=median)
```

***

### Plot the Grand Mean
Notes: Users with longer tenture have higher friend count with exception of older users about 80 and up

```{r Plot the Grand Mean}
ggplot(aes(x=age, y=friend_count), data=subset(pf, !is.na(year_joined))) +
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=mean) +
  geom_line(linetype=2, stat='summary', fun.y=mean)

#plotting the grand mean is a reminder that much of the data on this sample is about members of recent
#cohorts

```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure>=1), summary(friend_count/tenure))
# max is an outlier considering out data since out 3r quartie is 0.57 
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot(aes(x=tenure, y=friendships_initiated/tenure), data=subset(pf, !is.na(year_joined) & tenure>=1  )) +
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=mean)

#it apppears that people with more tenure typically initiated less friendships

```

***

### Bias-Variance Tradeoff Revisited
Notes: A lot of noise in the above grapth since we plot mean of y for every possible tenure x value

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

#graph with very high bias  byt much less variance
ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)
# as bin size increases we see less noise in the graph. our estimates of the mean are adjusted since we have 
#more data points for new values of the tenure

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))
# we still see that frienships initiated decline as tenure increases

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <-read.csv('yogurt.csv')
str(yo)

#Change the id from int to a factor
yo$id<-factor(yo$id)
str(yo)

qplot(data=yo, x=price) 
#discreteness, there appear to be prices where there are many observations and no observations
# in adjacent prices that makes snece because prices are set that applies to many consumers

qplot(data=yo, x=price, binwidth=10) 
#for very descrete data like this above histogram (with biwidth=10) is very biased model


```

***

### Number of Purchases
Notes: discreteness (notice that 75% percentile is the same as max), 20 discrete unique prices

```{r Number of Purchases}
summary(yo)
length(unique(yo$price))
table(yo$price) #<-- distribution of prices

yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
summary(yo$all.purchases)

#alternate version
yo$all.purchases <- yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry
```

***

### Prices over Time
Notes: discrentenes. Most housholds buy 1-2 yogurts at a time

```{r Prices over Time}
qplot(x=all.purchases, data=yo, binwidth =1)

ggplot(aes(x=time, y=price), data=yo ) + geom_jitter(alpha=1/4, shape=21)
#modal of the most common prices seem to increase over. WE also see some lower price points 
# scattered about the grapth() these maybe to sales or buyer using coupns that bring down prices of yogurt
```


***

### Sampling Observations
Notes: may be usefull to look into samll sample data in detail so we know what type of within and between household variation we are working with

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids <-sample(levels(yo$id),16)

ggplot(aes(x=time, y=price), data=subset(yo, id %in% sample.ids)) +
  facet_wrap(~ id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch=1)


#my plot
set.seed(999)
sample.ids <-sample(levels(yo$id),16)

ggplot(aes(x=time, y=price), data=subset(yo, id %in% sample.ids)) +
  facet_wrap(~ id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch=1)

```

***

### The Limits of Cross Sectional Data
Notes: If we have observations over time we can facet by primary unit ,case or individual  (household)

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
#install.packages('GGally') 
library(GGally)
theme_set(theme_minimal(20))

#set the seed for reproducible results
set.seed(1836)
pf_subset <- pf[ , c(2:15)] 
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset),100), ], axisLabels = 'internal')

```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")

#changing the colnames to produce a nicer plot
colnames(nci) <- c(1:64)
```

```{r}
#melt the data to long format
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

